Box 1 - TDT Curves

### 2023-09-24 ms/ms.qmd
Fixed equation

# Grab our data
   data <- read.csv(here::here("data", "drosi_data.csv"))
```

@@

```
```{r, posteriorsz}
# We'll now plot the posterior distributions for each species.Obviously our 11 species with known data will have a much tighter distribution. You can see that in the plot. But the imputations are our 'best" guess based on the strong phylogenetic signal in the data. We can also always propagate the uncertainty across any subsequent analysis (like for example our micro-climate models)
  mcmc_areas(z[,1:15],prob = 0.8)
```

@@

We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX. We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022. Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021. These data were integrated with data on lethal limits at the species-level. When the same species were sampled multiple times we retained within species variation in models. We then fit the following model in R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`)[@R2023] with the package *MCMCglmm* (vers. `r utils::packageVersion("MCMCglmm")`), which uses data augmentation during model fittng to estimate missing data [@Hadfield2010]:

@@

$$
p_{i,k} \sim N\left(0, \right) \\

sp_{i,k} \sim N\left(0, \right) \\

e_{i,k} \sim N\left(0,  \right) \\
$$

@@

We obtained data for `r` species (). Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`). While the model also suggested that there was within species variation, missing data and lack of replicate species populations make conclusions premature ($I^2_{species}$: $CT_{max}$ – `r posterior.mode(spp_ctmax)`, 95% CI: `r HPDinterval(spp_ctmax)[,"lower"]` to `r HPDinterval(spp_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(spp_z)`, 95% CI: `r HPDinterval(spp_z)[,"lower"]` to `r HPDinterval(spp_z)[,"upper"]`; CFLs – `r posterior.mode(spp_cfls)`, 95% CI: `r HPDinterval(spp_cfls)[,"lower"]` to `r HPDinterval(spp_cfls)[,"upper"]`).

>>>>

# Grab our data
   data <- read.csv(here::here("data", "drosi_data.csv"))

##################################
# Summary Data
##################################

sum <- data  %>% group_by(species)  %>% summarise(ctmax = mean(ctmax, na.rm = TRUE),
                                                   cfls = mean(cfts, na.rm = TRUE),
                                                   tdt_z = mean(cfts, na.rm = TRUE))

sum2 <- sapply(sum, function(x) sum(!is.na(x)))
```

@@

```{r, posteriorsz}
# We'll now plot the posterior distributions for each species.Obviously our 11 species with known data will have a much tighter distribution. You can see that in the plot. But the imputations are our 'best" guess based on the strong phylogenetic signal in the data. We can also always propagate the uncertainty across any subsequent analysis (like for example our micro-climate models)
  mcmc_areas(z[,1:15],prob = 0.8)
```

@@

We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX. We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022. Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021. These data were integrated with data on lethal limits at the species-level. Overall, we obtained data for `r length(unique(t$tip.label))` species ($N_{CT_{max}}$ = `r sum2[2]` species; $N_{z^{TDT}}$ = `r sum2[4]` species and $N_{CFLs}$ = `r sum2[3]` species). When the same species were sampled multiple times we retained within species variation in models. We then fit the following model in R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`)[@R2023] with the package *MCMCglmm* (vers. `r utils::packageVersion("MCMCglmm")`), which uses data augmentation during model fittng to estimate missing data [@Hadfield2010]:

@@

$$
\begin{aligned}
& p_{i,k} \sim N\left(0, \sigma_{p}^2 \textbf{A} \right) \\
& sp_{i,k} \sim N\left(0, \sigma_{sp}^2\textbf{I} \right) \\
& e_{i,k} \sim N\left(0,  \sigma_{e}^2\textbf{I}\right) \\
\end{aligned}
$$

@@

 Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`). While the model also suggested that there was within species variation, missing data and lack of replicate species populations make conclusions premature ($I^2_{species}$: $CT_{max}$ – `r posterior.mode(spp_ctmax)`, 95% CI: `r HPDinterval(spp_ctmax)[,"lower"]` to `r HPDinterval(spp_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(spp_z)`, 95% CI: `r HPDinterval(spp_z)[,"lower"]` to `r HPDinterval(spp_z)[,"upper"]`; CFLs – `r posterior.mode(spp_cfls)`, 95% CI: `r HPDinterval(spp_cfls)[,"lower"]` to `r HPDinterval(spp_cfls)[,"upper"]`).
